---
title: "Introduction to TPLthemeR"
author: "Heidi Wallace"
date: "`r format(Sys.time(), '%B %d, %Y, %H:%M')`"
output: 
  html_document:
    keep_md: true
    toc: true
    toc_depth: 3

---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 8, 
  fig.height = 6, 
  message = FALSE,
  warning = FALSE
)
```


### Installation & Setup

This is a brief guide to using the `TPLthemeR` package, which provides themes and colors for customizing ggplot visualizations using TPL's style guide. 


##### Install the package

The custom package can be installed to your R library directly from GitHub using the following commands in RStudio:

```{r, eval=FALSE}
install.packages("devtools") #devtools allows installation directly from github; only run this command if you do not already have devtools installed

devtools::install_github("heidiwallace/TPLthemeR") #install the package to your R library
```

```{r}
 #load the packages
library(TPLthemeR)
library(ggplot2)
library(tidyverse)
```

<br>

**Note**: If you run into trouble installing the `devtools` package, you may need to download Xcode (on Mac) or Rtools (on Windows) first. See <a href="https://www.r-project.org/nosvn/pandoc/devtools.html">here</a> for more details or let me know if I can help troubleshoot! 

<br>

##### Install the font to your device

Next, download the Open Sans font family from <a href="https://fonts.google.com/specimen/Open+Sans#standard-styles">the developer webpage</a>. Unzip the downloaded folder, then follow the instructions to import the new font into your system for <a href="https://www.pcworld.com/article/394991/how-to-install-fonts-in-windows-10.html"> Windows</a> or <a href="https://support.apple.com/guide/font-book/install-and-validate-fonts-fntbk1000/mac">Mac</a>. 

<br>

##### Register fonts with R

Before using the package, you'll also need to install the <a href="https://github.com/wch/extrafont">extrafonts</a> package to register the font being used in this package with R. Below are the commands to install the `extrafonts` package and register the fonts installed on your device with R. You only need to register the fonts once per device, and R will always have access to them. See the `extrafonts` github page for more details about this package.

```{r, eval=FALSE}
install.packages('extrafont')
library(extrafont)

font_import() #this command recognizes all the fonts loaded onto your computer and makes them accessible to R - may take a few minutes to run

loadfonts() #registers the imported fonts with R
```

<br>

##### Troubleshooting

There does seem to be an issue with another required package which caused the `font_import()` function to return the following message: "No FontName. Skipping." If this occurs on your device, you can <a href="https://stackoverflow.com/questions/61204259/how-can-i-resolve-the-no-font-name-issue-when-importing-fonts-into-r-using-ext">fix it</a> by first downgrading the Rttf2pt1 package (requires Xcode or Rtools) and then running `font_import()` again. Here are the commands:

```{r, eval=FALSE}
remotes::install_version("Rttf2pt1", version = "1.3.8")
extrafont::font_import()
extrafont::loadfonts()
```


<br><br>

### Plotting with custom theme

The package currently provides three custom themes that can be added to ggplot visualizations in the same way that `ggplot2`'s built-in themes are layered onto plots. The custom theme functions are `theme_tpl()`, `theme_tpl_legend_top()`, and `theme_tpl_legend_right()`.

`theme_tpl()` customizes the plot grid and fonts. It does not provide any customization for legends, though additional customization can be layered on manually using the typical `theme()` syntax. The additional themes called `theme_tpl_legend_top()`, and `theme_tpl_legend_right()` provide customized settings for legends placed either at the top of the plot (underneath the plot title) or to the right of the chart.

```{r}
#plot with default ggplot2 theme
p <- 
  diamonds %>% 
  ggplot(aes(cut, price)) +
  geom_col() +
  labs(title = "Prices of diamonds by cut",
       caption = "Here is a sample caption")

p
```

```{r}
#plot with custom TPL theme
p + theme_tpl()
```


<br><br>


Note also that the `theme_tpl()` function can be used in conjunction with typical `theme()` layers to adjust settings. If you wish to override any settings in `theme_tpl()`, `theme()` must follow the `theme_tpl()` layer.

```{r}
#plot with custom TPL theme & adjustments
p + 
  theme_tpl() +
  theme(axis.title.x = element_blank(),
        axis.text.y = element_text(size = 11),
        plot.caption = element_text(face = "italic"))
  
```

<br><br>


### Setting colors using plot layers

The package also provides four built-in functions called `scale_fill_discrete_tpl()`, `scale_fill_continuous_tpl()`, `scale_color_discrete_tpl()`, and `scale_color_continuous_tpl()`. These can be layered on to plots to provide branded coloring. These functions work just like the built-in `ggplot` functions `scale_fill_discrete()`,  `scale_fill_continuous()`, `scale_color_discrete()`, and `scale_color_continuous()`, and they accept the same arguments as the built-in functions. 

See `?scale_fill_continuous_tpl` and `?scale_color_continuous_tpl` for the two additional arguments that these functions require.


```{r}
#example with discrete colors

diamonds %>% 
  ggplot(aes(cut, price, fill = cut)) +
  geom_col() +
  labs(title = "Prices of diamonds by cut",
       caption = "Here is a sample caption") + 
  scale_fill_discrete_tpl() +
  theme_tpl() +
  theme(legend.position = "none",
        axis.title.x = element_blank())
```

<br><br>



```{r}
#example with continuous colors

g <- 
  ggplot(faithfuld, aes(waiting, eruptions, fill = density)) +
  geom_tile()

g + 
  theme_tpl_legend_right() +
  scale_fill_continuous_tpl('teal', reverse = T)

```

<br><br>

### Setting colors manually

TPL colors can also be called using the function `tpl_colors()`. This function creates a vector of named color hex codes. The colors and their codes are shown below.

```{r,echo=FALSE}
cols <- tpl_colors()
n_seq <- seq_along(cols)
image(n_seq, 1, as.matrix(n_seq), col = cols,
      xlab = "", ylab = "", xaxt = "n", yaxt = "n", bty = "n",
      main = "The People Lab Colors")
text(n_seq, 1, paste(names(cols), cols, sep = "\n"), col = "white", srt = 0)
```

<br><br>

A use case for this function is setting a single color or set of colors in a plot.

```{r}
#create a vector of color hex codes
colors <- tpl_colors()

#plot using a hex code from the vector
diamonds %>% 
  ggplot(aes(cut, price)) +
  geom_col(fill = colors[['orange']]) +
  labs(title = "Prices of diamonds by cut",
       caption = "Here is a sample caption") +
  theme_tpl()
```


### References

* <a href="http://rpubs.com/mclaire19/ggplot2-custom-themes">Learning to create custom themes in ggplot</a>

* <a href="https://drsimonj.svbtle.com/creating-corporate-colour-palettes-for-ggplot2">Creating corporate colour palettes for ggplot2</a>

* <a href="https://tinyheero.github.io/jekyll/update/2015/07/26/making-your-first-R-package.html">Making your first R package</a>
