---
title: "Introduction to TPLthemeR"
author: "Heidi Wallace"
date: "`r format(Sys.time(), '%B %d, %Y, %H:%M')`"
output: 
  html_document:
    keep_md: true
    toc: true
    toc_depth: 3

---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 8, 
  fig.height = 6, 
  message = FALSE,
  warning = FALSE
)
```


### Installation & Setup

This is a brief guide to using the `TPLthemeR` package, which provides themes and colors for customizing ggplot visualizations using TPL's style guide. 


##### Install the package

The custom package can be installed to your R library directly from GitHub using the following commands in RStudio:

```{r, eval=FALSE}
install.packages("devtools") #devtools allows installation directly from github; only run this command if you do not already have devtools installed

devtools::install_github("heidiwallace/TPLthemeR") #install the package to your R library
```

<br>

**Note**: If you run into trouble installing the `devtools` package, you may need to download Xcode (on Mac) or Rtools (on Windows) first, then re-run these commands. See <a href="https://www.r-project.org/nosvn/pandoc/devtools.html">here</a> for more details or let me know if I can help troubleshoot! 


<br><br>

### Plotting with custom theme

The package currently provides three custom themes that can be added to ggplot visualizations in the same way that `ggplot2`'s built-in themes are layered onto plots. The custom theme functions are `theme_tpl()`, `theme_tpl_legend_top()`, and `theme_tpl_legend_right()`.

`theme_tpl()` customizes the plot grid and fonts. It does not provide any customization for legends, though additional customization can be layered on manually using the typical `theme()` syntax. The additional themes called `theme_tpl_legend_top()`, and `theme_tpl_legend_right()` provide customized settings for legends placed either at the top of the plot (underneath the plot title) or to the right of the chart. These are demonstrated in a later section.

```{r}
 #load the packages
library(TPLthemeR)
library(ggplot2)
library(tidyverse)
```

```{r}
#plot with default ggplot2 theme
p <- 
  diamonds %>% 
  ggplot(aes(cut, price)) +
  geom_col() +
  labs(title = "Prices of diamonds by cut",
       caption = "Here is a sample caption")

p
```

```{r}
#plot with custom TPL theme
p + theme_tpl()
```


<br><br>


Note also that the `theme_tpl()` function can be used in conjunction with typical `theme()` layers to adjust settings. If you wish to override any settings in `theme_tpl()`, `theme()` must follow the `theme_tpl()` layer.

```{r}
#plot with custom TPL theme & adjustments
p + 
  theme_tpl() +
  theme(axis.title.x = element_blank(),
        axis.text.y = element_text(size = 11),
        plot.caption = element_text(face = "italic"))
  
```

<br><br>


### Setting colors using plot layers

The package also provides four built-in functions called `scale_fill_discrete_tpl()`, `scale_fill_continuous_tpl()`, `scale_color_discrete_tpl()`, and `scale_color_continuous_tpl()`. These can be layered on to plots to provide branded coloring. These functions work just like the built-in `ggplot` functions `scale_fill_discrete()`,  `scale_fill_continuous()`, `scale_color_discrete()`, and `scale_color_continuous()`, and they accept the same arguments as the built-in functions. 

See `?scale_fill_continuous_tpl` and `?scale_color_continuous_tpl` for the two additional arguments that these functions require.


```{r}
#example with discrete colors

diamonds %>% 
  ggplot(aes(cut, price, fill = cut)) +
  geom_col() +
  labs(title = "Prices of diamonds by cut",
       caption = "Here is a sample caption") + 
  scale_fill_discrete_tpl() +
  theme_tpl() +
  theme(legend.position = "none",
        axis.title.x = element_blank())
```

<br>

It's important to note that there only five colors in the custom palette, so if you have more than five discrete categories, ggplot will revert to its default colors to accommodate the additional categories. 

<br><br>



```{r}
#example with continuous colors

g <- 
  ggplot(faithfuld, aes(waiting, eruptions, fill = density)) +
  geom_tile()

g + 
  theme_tpl_legend_right() +
  scale_fill_continuous_tpl('teal', reverse = T)

```


<br><br>

### Using legends

`theme_tpl_legend_top()` and `theme_tpl_legend_right()` are layers to that apply the custom TPL theme to legends. Use `theme_tpl_legend_top()` to set the legend underneath the title and `theme_tpl_legend_right()` to set the legend to the right of the plot.

```{r}
#a plot with right legend
diamonds %>% 
  group_by(color, cut) %>% 
  summarise(price = mean(price)) %>% 
  ggplot(aes(color, price, fill = cut)) +
  geom_col(position = "dodge") +
  labs(title = "Prices of diamonds by color and cut",
       caption = "Here is a sample caption") +
  theme_tpl_legend_right() +
  scale_fill_discrete_tpl()
```

<br><br>

```{r}
#a plot with top legend
diamonds %>% 
  group_by(color, cut) %>% 
  summarise(price = mean(price)) %>% 
  ggplot(aes(color, price, fill = cut)) +
  geom_col(position = "dodge") +
  labs(title = "Prices of diamonds by color and cut",
       caption = "Here is a sample caption") +
  theme_tpl_legend_top() +
  scale_fill_discrete_tpl()

ggsave("test.png")
```


<br><br>

### Setting colors manually

TPL colors can also be called using the function `tpl_colors()`. This function creates a vector of named color hex codes. The colors and their codes are shown below.

```{r,echo=FALSE}
cols <- tpl_colors()
n_seq <- seq_along(cols)
image(n_seq, 1, as.matrix(n_seq), col = cols,
      xlab = "", ylab = "", xaxt = "n", yaxt = "n", bty = "n",
      main = "The People Lab Colors")
text(n_seq, 1, paste(names(cols), cols, sep = "\n"), col = "white", srt = 0)
```

<br><br>

A single color's hex code can be called using the name.

```{r}
#create a vector of color hex codes
colors <- tpl_colors()

colors[['darkblue']]
```


<br><br>

A use case for this function is setting a single color or set of colors in a plot.

```{r}
#plot using a hex code from the vector
diamonds %>% 
  ggplot(aes(cut, price)) +
  geom_col(fill = colors[['orange']]) +
  labs(title = "Prices of diamonds by cut",
       caption = "Here is a sample caption") +
  theme_tpl()
```


<br><br>

### References

* <a href="http://rpubs.com/mclaire19/ggplot2-custom-themes">Learning to create custom themes in ggplot</a>

* <a href="https://drsimonj.svbtle.com/creating-corporate-colour-palettes-for-ggplot2">Creating corporate colour palettes for ggplot2</a>

* <a href="https://tinyheero.github.io/jekyll/update/2015/07/26/making-your-first-R-package.html">Making your first R package</a>

* 
