---
title: "Introduction"
date: "`r format(Sys.time(), '%B %d, %Y, %H:%M')`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(TPLthemeR)
library(ggplot2)
library(tidyverse)
```


This is a brief guide to using the `TPLthemeR` package, which provides themes and colors for customizing ggplot visualizations using TPL's style guide. 

The package currently provides three custom themes that can be added to ggplot visualizations in the same way that `ggplot2`'s built-in themes are layered onto plots.

`theme_tpl()` customizes the plot grid and fonts. It does not provide any customization for legends.

```{r}
p <- 
  diamonds %>% 
  ggplot(aes(cut, price)) +
  geom_col() +
  labs(title = "Prices of diamonds by cut",
       caption = "Here is a sample caption")

p + theme_tpl()
```


The package also provides two built-in functions called `scale_fill_discrete_tpl()` and `scale_fill_continuous_tpl()` that can be layered on to provide branded coloring. These functions work just like the built-in `ggplot` functions `scale_fill_discrete()` and `scale_fill_continuous()` and accept the same arguments. See `?scale_fill_continuous_tpl` for the two additional arguments that this function requires.

Note also that the `theme_tpl()` function can be used in conjunction with typical `theme()` layers. If wishing to override settings in `theme_tpl()` using `theme()`, `theme()` must follow `theme_tpl()`.

```{r}
diamonds %>% 
  ggplot(aes(cut, price, fill = cut)) +
  geom_col() +
  labs(title = "Prices of diamonds by cut",
       caption = "Here is a sample caption") + 
  scale_fill_discrete_tpl() +
  theme_tpl() +
  theme(legend.position = "none")
```

```{r}
g <- 
  ggplot(faithfuld, aes(waiting, eruptions, fill = density)) +
  geom_tile()

g + scale_fill_continuous_tpl('teal', reverse = T)

```




TPL colors can also be called by the function `tpl_colors()`. This function creates a vector of named color hex codes. The colors and their codes are shown below.

```{r,echo=FALSE}
cols <- tpl_colors()
n_seq <- seq_along(cols)
image(n_seq, 1, as.matrix(n_seq), col = cols,
      xlab = "", ylab = "", xaxt = "n", yaxt = "n", bty = "n",
      main = "The People Lab Colors")
text(n_seq, 1, cols, col = "white", srt = 90)
```


A use case for this function is setting a single color or set of colors in a plot.

```{r}
colors <- tpl_colors()

diamonds %>% 
  ggplot(aes(cut, price)) +
  geom_col(fill = colors[['orange']]) +
  labs(title = "Prices of diamonds by cut",
       caption = "Here is a sample caption") +
  theme_tpl()
```


